# 04 - Classification {.unnumbered}

<style>

h3, h4 {
    position: absolute;
    width: 1px;
    height: 1px;
    padding: 0;
    margin: -1px;
    overflow: hidden;
    clip: rect(0,0,0,0);
    border: 0;
}

</style>

## Libraries

```{r}
library(tidymodels)
library(data.table)
library(discrim)
library(klaR)
library(poissonreg)
```

## Custom functions

In this section I place the functions that were used in many code chunks across the chapter.

```{r}


```


## Conceptual

### 1

**1.** Using basic statistical properties of the variance, as well as single-variable calculus, derive (5.6). In other words, prove that $\alpha$ given by (5.6) does indeed minimize $\text{Var}(\alpha X + (1-\alpha)Y)$.

$$
\begin{split}
\text{Var}(\alpha X + (1-\alpha)Y) & =  \alpha^2      \text{Var}(X)+
                                        (1-\alpha)^2  \text{Var}(Y) +
                                2 \alpha (1-\alpha)   \text{Cov}(X,Y) \\
    & = \alpha^2      \text{Var}(X)+
        \text{Var}(Y)
        -\alpha^2 \text{Var}(Y) +
        2 \alpha (1-\alpha)   \text{Cov}(X,Y) \\
\end{split}
$$


